---
title: "Radial Visualization"
author: "Catherine Williams"
date: "July 6, 2019"
output:
  html_document:
    hightlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_depth: 2
    toc_float: yes
---
# Challenge Description

This month, the challenge is to create and share a radial data visualization. Any circular view of data is welcome, but the data and scenario should lend itself well to this.

# Dataset

Data came from [Kaggle](https://www.kaggle.com/selfishgene/historical-hourly-weather-data#weather_description.csv) and includes historical hourly weather data for 36 different cities from 2012-2017. However, only Seattle data will be used in this project.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
setwd("~/../Google Drive/Data Analysis/SWD Challenges/07-2019/")
```

# Setup

## Load Libraries

```{r library}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse","visdat","lubridate","packcircles","ggforce")
```

## Import Data

```{r import}
df <- read_csv("weather_description.csv")
```

# Data Wrangling

- There are 45,253 observations.
- Each city has its own column with a character description of the weather pattern
- The first row is NA for all cities.

## View Data

```{r view}
glimpse(df)
head(df)

levels(as.factor(df$Seattle))
summary(df$datetime)
```

## Filter Data

Data before December 2012 was removed in order to have an equal number of each month so that Winter will not show skewed data.

```{r clean}
df <- df %>% select(datetime,Seattle)
df <- df %>% filter(datetime > "2012-11-30") 
```

## Missing Values

There are no missing values.

```{r missing}
#Visualize missing values
vis_miss(df, sort_miss=TRUE)
```

## Prepare Dataframe

```{r wrangle}
df <- df %>% transmute(month = month(datetime),
                      season = case_when(month %in% c(12,1,2) ~ "Winter",
                                       month %in% c(3,4,5) ~ "Spring",
                                       month %in% c(6,7,8) ~ "Summer",
                                       month %in% c(9,10,11) ~ "Fall"),
                      weather = case_when(str_detect(Seattle, "rain|drizzle") ~ "rain",
                                          str_detect(Seattle, "mist|fog") ~ "fog",
                                          str_detect(Seattle, "clear") ~ "clear",
                                          str_detect(Seattle, "squalls") ~ "squalls",
                                          str_detect(Seattle, "cloud") ~ "clouds",
                                          str_detect(Seattle, "smoke|haze") ~ "haze",
                                          str_detect(Seattle, "snow") ~ "snow",
                                          str_detect(Seattle, "thunderstorm") ~ "thunder"))

# aster chart
df_aster <- df %>% group_by(weather) %>% summarize(count=n()) %>% arrange(count) %>% ungroup()
df_aster <- df_aster %>% mutate_at(vars(weather), as.factor)
df_aster$weather <- fct_reorder(df_aster$weather, df_aster$count)
df_aster <- df_aster %>% filter(count > 20)

# circle chart
df_circle <- df %>% group_by(season,weather) %>% summarize(count=n()) %>% arrange(count) %>% ungroup()
df_circle <- df_circle %>% mutate_if(is.character, as.factor)
df_circle <- df_circle %>% filter(count > 20)
```

# Visualizations
## Aster Chart

```{r aster}
df_aster %>% ggplot(aes(weather,sqrt(count),fill=weather))+
      geom_col()+
      geom_text(aes(y=50,label=weather))+
      coord_polar()+
      ylim(-10,125)+
      labs(title="Seattle Weather Occurrences",subtitle="Measured from 2012 - 2017",fill="Weather Type")+
      theme_minimal()+
      theme(axis.title = element_blank(),
            axis.text = element_blank(),
            plot.title = element_text(hjust=0.5),
            plot.subtitle = element_text(hjust=0.5),
            legend.position = "none")
```

## Circle Packed Chart

```{r}
set.seed(333)
df_packs <- df_circle$count %>% circleRepelLayout() %>% .$layout

df_packs <- bind_cols(df_circle, df_packs) %>% select(-count)

# set colors
colors <- c("#BFD1E5","#8CC7A1","#C1666B","#D4B483") %>% set_names(nm = c("Winter","Spring","Summer","Fall"))

# main circle plot
p <- df_packs %>% ggplot() +
  geom_circle(aes(x0=x, y0=y, r=radius, fill=season, color=season), size=0.1, show.legend=FALSE) +
  geom_text(aes(x=x, y=y, label=weather, size=radius*0.5), show.legend=FALSE) +
  scale_fill_manual(values=colors) +
  scale_color_manual(values=colors) +
  coord_fixed() +
  ggtitle("Seattle Weather Occurrences")+
  theme_void() +
  theme(text = element_text(family="courier"),
        plot.title = element_text(face="bold", hjust=0.5, size=14),
        plot.background = element_rect(fill="snow", color="black"))

# make legend maually
legend <-
  tibble(y=rep(1, each=4), x=rep(1:4, times=1), fill=colors %>% names(), r=0.1) %>%
  ggplot() +
  geom_text(aes(x=x, y=y-.2, label=fill), size=2.25, vjust=1, family="courier", fontface="bold") +
  geom_circle(aes(x0=x, y0=y, fill=fill, color=fill, r=r), show.legend=FALSE) +
  scale_fill_manual(values=colors) +
  scale_colour_manual(values=colors) +
  coord_fixed() +
  theme_void() +
  lims(x=c(.5,6), y=c(0.5, 2.3))

leg_text <- tibble(x=0.16, y=0.17, label="Season") %>% 
  ggplot()+
  geom_text(aes(x=x, y=y, label=label), family="courier", fontface="bold", size=3.75, show.legend=FALSE)+
  coord_fixed()+
  lims(x=c(-4,5), y=c(-1, 1))+
  theme_void()

# combine plot with legend
p + annotation_custom(
      grob = ggplotGrob(legend),
      xmin = -190,
      xmax = -40,
      ymin = -125,
      ymax = -75) +
    annotation_custom(
      grob = ggplotGrob(leg_text),
      xmin = -165,
      xmax = -100,
      ymin = -110,
      ymax = -90)

# save final plot
ggsave("seattle-weather.png", width=6.1)
```