---
title: "House Prices"
author: "Catherine Williams"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_fragment:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

directory <- "C:/Users/Cat/Google Drive/Data Analysis/Kaggle Competitions/House Prices - Advanced Regression"

if(!getwd() == directory) {
  setwd(directory)
}
```

## Load Libraries
```{r library}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse","modeest","zoo","corrplot","reshape2","naniar","qqplotr","fastDummies","FNN","Boruta","caret","performance","caretEnsemble","doParallel")

theme_set(theme_classic())
```

Import the data that Kaggle provided.

```{r import, message=FALSE}
#import data frames
df_train <- read_csv("train.csv")
df_test <- read_csv("test.csv")
```
# Exploratory Data Analysis

Review data to understand it.

## View Data
```{r review}
#review data frame
df_train %>% glimpse()
df_test %>% glimpse()

df_train$SalePrice %>% summary()
df_train %>% count(Neighborhood, sort=TRUE)
```

Combine data for cleaning so that the same manipulations are applied to everything.

```{r combine}
#combine data for cleaning and viewing
df <- bind_rows(df_train, df_test)

df %>% glimpse()

#view unique values
sapply(df, function(x) n_distinct(x)) %>% sort()
```

## Missing Values

```{r na, warning=FALSE}
#Visualize missing values
gg_miss_var(df[,colSums(is.na(df)) > 0], show_pct=TRUE) + labs(title="Missing Values")

#see count of missing values
na_values <- function(df){
  na <- colSums(is.na(df)) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df)
```

## Correlations

Find correlated variables

```{r corr}
corr_simple <- function(data=df,drop="SalePrice",sig=0.5){
  #convert data to numeric in order to run correlations
  df_cor <- data %>% mutate_if(is.character, as.factor) %>% select(-drop)
  df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)

  #run a correlation and since there are so many variables, drop the insignificant ones
  corr <- cor(df_cor)
  corr[lower.tri(corr,diag=TRUE)] <- NA  #Prepare to drop duplicates and correlations of 1
  corr[corr == 1] <- NA #drop perfect correlations
  corr <- as.data.frame(as.table(corr)) #Turn into a 3-column table
  corr <- na.omit(corr) #remove the NA values from above
  corr <- subset(corr, abs(Freq) > sig) #select significant values
  corr <- corr[order(-abs(corr$Freq)),] #Sort by highest correlation
  print(corr)

  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  
  #plot correlations visually
  corrplot::corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}

corr_simple()
```

# Data Wrangling

## Fix Missing Values

Replace missing values based on appropriate logic.

PoolQC, MiscFeature, Alley and Fence have a lot of missing values, however, in this case NA means the features do not exist. These will be changed to character values of NA.

```{r}
#the following features do not necessarily exist. These can be changed to "NA"
df$MiscFeature[is.na(df$MiscFeature)] <- "NA"
df$Alley[is.na(df$Alley)] <- "NA"
df$Fence[is.na(df$Fence)] <- "NA"
df$GarageType[is.na(df$GarageType)] <- "NA"
df$MasVnrArea[is.na(df$MasVnrArea)] <- 0

#the following features do not necessarily exist but are dependent on values in other columns. They will be changed to "NA" based on other column values
df$PoolQC[df$PoolArea == 0] <- "NA"
df$FireplaceQu[df$Fireplaces == 0] <- "NA"
df$GarageYrBlt[df$GarageType == "NA"] <- "NA"
df$GarageFinish[df$GarageType == "NA"] <- "NA"
df$GarageQual[df$GarageType == "NA"] <- "NA"
df$GarageCond[df$GarageType == "NA"] <- "NA"
df$MasVnrType[df$MasVnrArea == 0] <- "None"

#there are a lot of basement fields so we want to make sure that this is handled properly
#first make a basement variable for the missing variables to depend on, based on other variables. 
#Since none of the below variables by itself can definitively say whether a basement exists or not, we see where the sums of all of them equal 0 and decide then that a basement does not exist
df <-  df %>% mutate(Bsmt = if_else(TotalBsmtSF + BsmtFullBath + BsmtHalfBath + BsmtUnfSF + BsmtFinSF1 + BsmtFinSF2 >= 1, TRUE, FALSE))

df$Bsmt[df$Bsmt %in% NA] <- FALSE

df$BsmtCond[df$Bsmt == FALSE] <- "NA"
df$BsmtExposure[df$Bsmt == FALSE] <- "NA"
df$BsmtQual[df$Bsmt == FALSE] <- "NA"
df$BsmtFinType2[df$Bsmt == FALSE] <- "NA"
df$BsmtFinType1[df$Bsmt == FALSE] <- "NA"
df$BsmtFullBath[df$Bsmt == FALSE] <- 0
df$BsmtHalfBath[df$Bsmt == FALSE] <- 0
df$BsmtFinSF1[df$Bsmt == FALSE] <- 0
df$BsmtFinSF2[df$Bsmt == FALSE] <- 0
df$BsmtUnfSF[df$Bsmt == FALSE] <- 0
df$TotalBsmtSF[df$Bsmt == FALSE] <- 0

df <- df %>% mutate(BsmtCond = replace_na(BsmtCond, mfv(BsmtCond, na.rm=T)),
                    BsmtExposure = replace_na(BsmtExposure, mfv(BsmtExposure, na.rm=T)),
                    BsmtQual = replace_na(BsmtQual, mfv(BsmtQual, na.rm=T)),
                    BsmtFinType2 = replace_na(BsmtFinType2, mfv(BsmtFinType2, na.rm=T)))

#these features should be determined by central tendency statistics, in this case- mean and mode
df <- df %>% mutate(GarageArea = na.aggregate(GarageArea,GarageType),
                    GarageCars = na.aggregate(GarageCars,GarageType) %>% round(),
                    GarageQual = replace_na(GarageQual, mfv(GarageQual, na.rm=T)),
                    GarageCond = replace_na(GarageCond, mfv(GarageCond, na.rm=T)),
                    PoolQC = replace_na(PoolQC, mfv(PoolQC, na.rm=T)),
                    Utilities = replace_na(Utilities, mfv(Utilities, na.rm=T)),
                    Exterior1st = replace_na(Exterior1st, mfv1(Exterior1st, na.rm=T)),
                    Exterior2nd = replace_na(Exterior2nd, mfv1(Exterior2nd, na.rm=T)))

#find mode based on a logical group
df <-df %>% group_by(YearRemodAdd) %>% mutate(GarageYrBlt = replace_na(GarageYrBlt, mfv1(GarageYrBlt, na.rm=T))) %>% ungroup()

df <-df %>% group_by(GarageYrBlt) %>% mutate(GarageFinish = replace_na(GarageFinish, mfv1(GarageFinish, na.rm=T))) %>% ungroup()

df <-df %>% group_by(OverallQual) %>% mutate(KitchenQual = replace_na(KitchenQual, mfv(KitchenQual, na.rm=T))) %>% ungroup()

df <- df %>% group_by(OverallCond)%>% mutate(Functional = replace_na(Functional, mfv(Functional, na.rm=T))) %>% ungroup()

df <- df %>% group_by(Neighborhood)%>% mutate(MSZoning = replace_na(MSZoning, mfv(MSZoning, na.rm=T)),
                                              MasVnrType = replace_na(MasVnrType, mfv(MasVnrType, na.rm=T)),
                                              LotFrontage = na.aggregate(LotFrontage)) %>% ungroup()

df <- df %>% group_by(YearBuilt)%>% mutate(Electrical = replace_na(Electrical, mfv1(Electrical, na.rm=T))) %>% ungroup()

df <- df %>% group_by(SaleCondition)%>% mutate(SaleType = replace_na(SaleType, mfv(SaleType, na.rm=T))) %>% ungroup()

#check for final na_values
na_values(df)
```

## Strings

Correct issues with strings.

```{r strings}
#fix strings that do not match the legend and/or have typos
df <- df %>% mutate(Exterior2nd = case_when(Exterior2nd %in% "Brk Cmn" ~ "BrkComm",
                                            Exterior2nd %in% "Wd Shng" ~ "WdShing",
                                            Exterior2nd %in% "CmentBd" ~ "CemntBd",
                                            TRUE ~ Exterior2nd)) 
```

## Data Types

Change variable types to more appropriate types.

```{r clean, message=FALSE, warning=FALSE}
#change variable types & feature engineering
df %>% glimpse()

df <- df %>% mutate_if(is.character, as.factor)

df <- df %>% mutate(Id = as.factor(Id),
                    MSSubClass = as.factor(MSSubClass),
                    MoSold = as.factor(MoSold),
                    CentralAir = if_else(CentralAir=="Y",TRUE,FALSE) %>% as.logical(),
                    SsnPorch = `3SsnPorch`,
                    TotalSF = `1stFlrSF`+`2ndFlrSF`+BsmtFinSF1+BsmtFinSF2,
                    NumBath = FullBath+(HalfBath*0.5)+BsmtFullBath+(BsmtHalfBath*0.5),
                    Garage = GarageType!="NA" & GarageType!="CarPort",
                    CarPort = GarageType=="CarPort",
                    Residential = str_detect(MSZoning,"R.") | MSZoning=="FV",
                    Sewer = if_else(Utilities=="AllPub",TRUE,FALSE) %>% as.logical()) %>% 
  select(-`1stFlrSF`, -`2ndFlrSF`, -`3SsnPorch`,-BsmtFinSF1,-BsmtFinSF2,- Utilities)

df <- df %>% mutate_at(vars(BedroomAbvGr,KitchenAbvGr,TotRmsAbvGrd,Fireplaces,GarageCars,YearBuilt,YearRemodAdd),as.integer)

df <- df %>% mutate_at(vars(LandSlope,OverallQual,OverallCond,ExterQual,ExterCond,BsmtQual,BsmtCond,BsmtExposure,BsmtFinType1,BsmtFinType2,HeatingQC,KitchenQual,Functional,FireplaceQu,GarageFinish,GarageQual,GarageCond,PavedDrive,PoolQC), as.ordered)

cols <- df %>% select_if(., is.ordered) %>% names()

#view levels to make sure they match up properly
view_levels <- function(cols){
  for(col in cols){
  lev <- df[[col]] %>% levels()
  cat(col, ": ", lev, "\n")
  }
}

view_levels(cols)

#change levels. If they match these values, they will change. If not, they will get ignored. The majority of ordered variables match these values.
for(col in cols){
  df[[col]] <- fct_relevel(df[[col]],c("NA","Po","Fa","TA","Gd","Ex"))
}

df$LandSlope <- fct_rev(df$LandSlope)
df$BsmtExposure <- fct_relevel(df$BsmtExposure,c("NA","No","Mn","Av","Gd"))
df$BsmtFinType1 <- fct_relevel(df$BsmtFinType1,c("NA","Unf","LwQ","Rec","BLQ","ALQ","GLQ"))
df$BsmtFinType2 <- fct_relevel(df$BsmtFinType2,c("NA","Unf","LwQ","Rec","BLQ","ALQ","GLQ"))
df$Functional <- fct_relevel(df$Functional,c("Sal","Sev","Maj2","Maj1","Mod","Min2","Min1","Typ"))
df$GarageFinish <- fct_relevel(df$GarageFinish,c("NA","Unf","RFn","Fin"))


#check more factor levels
cols <- df %>% select(Exterior1st,Exterior2nd,Condition1,Condition2,BsmtFinType1,BsmtFinType2) %>% names()
view_levels(cols)

df$Exterior1st <- factor(df$Exterior1st, levels=levels(df$Exterior2nd))
df$Condition2 <- factor(df$Condition2,levels=levels(df$Condition1))
```

## Feature Engineering

Remove the following columns:

- All quality and condition columns except for 'Overall' which should cover everything.
- GarageCars - highly correlated with GarageArea. It is pretty obvious that if more cars can fit, the area is larger.
- TotalRmsAbvGrd, GrLivArea, and TotalBsmtSF are highly correlated with TotalSF. TotalSF seems to be a more important measure.
- FullBath and BsmtFullBath are largely duplicate information with NumBath.
- GarageYrBlt is highly correlated with YearBuilt and YearRemodAdd. These other features describe the necessary information.
- MSSubClass is highly correlated with BldgType. It also appears to contain information about the year built, even though that does not show up as highly correlated.
- YearBuilt - highly correlated with YearRemodAdd which is the same date if there was no remodel. YearBuilt is also highly correlated with many other features so this seems duplicated and somewhat irrelevant compared to remodel dates.

Information changed:

- Exterior2nd and BsmtFinType2 had duplicate values if there were not two different types. Changing these fields to "NA" or "None" is more accurate and reduced the correlations.
- Condition2 also had duplicate values if there were not two different types, but this made more sense to create new variables summarizing the important qualities of Condition1 and Condition2.

```{r}
#Remove unneeded columns
df <- df %>% select(-ExterQual,-ExterCond,-BsmtQual,-BsmtCond,-HeatingQC,-KitchenQual,-FireplaceQu,-GarageQual,-GarageCond,-PoolQC,-GarageArea,-TotRmsAbvGrd,-GrLivArea,-FullBath,-GarageYrBlt,-MSSubClass,-TotalBsmtSF,-GarageFinish,-YearBuilt,-BsmtFullBath,-HalfBath)

#Remove unnecessary duplicate information
df <- df %>% mutate_at(vars(Exterior1st,Exterior2nd,BsmtFinType1,BsmtFinType2),as.character)

df <- df %>% mutate(Railroad = if_else(str_detect(Condition1, "RR"), TRUE,
                                      if_else(str_detect(Condition2, "RR"), TRUE, FALSE)),
                    PositiveFeature = if_else(str_detect(Condition1, "Pos"), TRUE,
                                              if_else(str_detect(Condition2, "Pos"), TRUE, FALSE)),
                    BusyStreet = if_else(str_detect(Condition1, "Artery|Feedr"), TRUE,
                                          if_else(str_detect(Condition2, "Artery|Feedr"), TRUE, FALSE)),
                    Exterior2nd = case_when(Exterior2nd == Exterior1st ~ "NA",
                                            TRUE ~ Exterior2nd),
                    BsmtFinType2 = case_when(BsmtFinType2 == BsmtFinType1 ~ "None",
                                            TRUE ~ BsmtFinType2),
                    SchoolDist = case_when(Neighborhood %in% c("Blmngtn","BrDale","NoRidge","NPkVill","NridgHt",
                                                                "NAmes","NWAmes","Somerst","StoneBr","Veenker") ~ "Fellows",
                                            Neighborhood %in% c("Blueste","ClearCr","Sawyer","SawyerW") ~ "Sawyer",
                                            Neighborhood %in% c("BrkSide","Crawfor","Names","OldTown") ~ "Meeker",
                                            Neighborhood %in% c("CollgCr","Edwards","IDOTRR","SWISU") ~ "Edwards",
                                            Neighborhood %in% "Gilbert" ~ "Gilbert",
                                            Neighborhood %in% c("MeadowV","Mitchel","Timber") ~ "Mitchell"),
                    OverallQC = as.integer(OverallQual) + as.integer(OverallCond),
                    OverallQC = as.ordered(OverallQC))

df$GarageCars[df$Garage == FALSE] <- 0

df <- df %>% mutate_at(vars(Exterior1st,Exterior2nd,BsmtFinType1,BsmtFinType2),as.factor) %>% 
  select(-Garage,-Condition1,-Condition2,-OverallQual,-OverallCond)
```

## Check Correlations Again

TotalSF and NumBath are highly correlated but it does not make sense to drop either one or to change them.

```{r corr 2}
corr_simple(data=df,sig=0.3)
```

# More Exploratory Data Analysis

- Although BsmtHalfBath did not show up highly correlated to anything, it does not make sense to keep in the model.
- LotFrontage, LotArea, BedroomAbvGr, TotalSF, NumBath appear to have extreme outliers.

## Variable Relationships

```{r relationship, warning=FALSE, out.width="50%"}
plot_relationships <- function(x_col, y_col="SalePrice", data=df){
  if(is.numeric(data[[x_col]])){
      scat <- ggplot(data, mapping=aes_string(x_col, y_col))+
        geom_jitter(alpha=0.5)+
        geom_smooth(method="loess", se=FALSE)+
        geom_smooth(method="lm", se=FALSE, color="red")+
        labs(title=str_c("Sale Price vs ", x_col))
      scat %>% print()
      }
  else{
    box <- ggplot(data, aes_string(x_col, y_col))+
      geom_boxplot()+
      labs(title=str_c("Sale Price by ", x_col))
    box %>% print()
    }
}

cols <- df %>% select(-1,-SalePrice) %>% names()
cols %>% walk(plot_relationships)
```

## Normality Check

```{r normality, warning=FALSE, out.width="50%"}
plot_normality <- function(x_col, data=df){
  if(is.numeric(data[[x_col]])){
      #histogram
      h <- hist(data[[x_col]], breaks = "FD", plot = FALSE) #histogram with Freedman-Diaconis rule for binwidth
      hist <- ggplot(data, aes_string(x_col))+
        geom_histogram(aes(y = ..density..), breaks = h$breaks, alpha = 0.3, col = "white")+
        geom_density(size = 1) +
        labs(title=str_c("Histogram and density for ", x_col))
      hist %>% print()
      #qqplot
      qq <- ggplot(data, aes_string(sample=x_col))+
        stat_qq_band(fill="#ffe6e6",detrend=TRUE)+
        stat_qq_line(color="#ff6666", size=1,detrend=TRUE)+
        stat_qq_point(detrend=TRUE)+
        labs(title=str_c("Quantile-quantile Normal plot of ", x_col), x="Theoretical Quantiles",y="Sample Quantiles")
      qq %>% print()
      #statistical tests
      sw <- shapiro.test(data[[x_col]])
      cat(x_col)
      print(sw)
  }
}

cols <- df %>% select(-1) %>% names()
cols %>% walk(plot_normality)
```

# More Data Wrangling

```{r drop}
#Remove unnecessary column
df <- df %>% select(-BsmtHalfBath)
```

## Transformations

Transform variables to make them more normally distributed.

```{r transform, warning=FALSE, out.width="50%"}
df <- df %>% 
   mutate(SalePrice.log = log(SalePrice),
          LotFrontage.sqr = sqrt(LotFrontage),
          LotArea.log = log(LotArea),
          MasVnrArea.sqr = sqrt(MasVnrArea),
          BsmtUnfSF.sqr = sqrt(BsmtUnfSF),
          WoodDeckSF.sqr = sqrt(WoodDeckSF),
          OpenPorchSF.sqr = sqrt(OpenPorchSF),
          EnclosedPorch.sqr = sqrt(EnclosedPorch),
          ScreenPorch.sqr = sqrt(ScreenPorch),
          MiscVal.log = log1p(MiscVal),
          TotalSF.log = log(TotalSF),
          NumBath.sqr = sqrt(NumBath)) %>% select(-SalePrice,-LotFrontage,-LotArea,-MasVnrArea,-BsmtUnfSF,-OpenPorchSF,-EnclosedPorch,-ScreenPorch,-MiscVal,-TotalSF,-NumBath)

cols <- df %>% select(51:62) %>% names()

for(col in cols){
 plot_normality(x_col=col)
  }
```
## View Relationships Again

```{r, warning=FALSE}
cols <- df %>% select(-1,-SalePrice.log) %>% names()
cols %>% walk(plot_relationships, y_col="SalePrice.log", data=df)
```

## Standardize Data

```{r norm}
df <- df %>% mutate_if(is.numeric, scale)

df %>% head()
```

## Training & Testing Sets

Recreate original testing dataframe as well as a training and validation set.

```{r dummy}
# set up new dataframe to turn factors into binary dummy variables for knn model
# df_dummy <- df %>% mutate_if(is.ordered, as.character) 
# df_dummy <- df_dummy %>% mutate_if(is.character, as.factor)
# 
# df_dummy <- df_dummy %>% select(-Id) %>% dummy_cols()
# df_dummy <- df_dummy %>% select_if(is.numeric)
# 
# df_dummy %>% str()
```

```{r test train}
#break training set into training and test sets
test <- df %>% filter(is.na(SalePrice.log))

train <- df %>% filter(!is.na(SalePrice.log))
```

## Feature Selection

The Boruta package will be used to decide which features are important. This is useful to reduce noise and speed up model training.

```{r boruta}
set.seed(123)
boruta_res <- Boruta(SalePrice.log ~. -Id, data = train, doTrace = 1)

print(boruta_res)
```

The following plot shows the relative importance of each candidate explanatory attribute. 

- Green indicates the attributes that are relevant to prediction.  
- Red is for attributes that are not relevant.  
- Yellow indicates attributes that may or may not be relevant to predicting the response variable.
- Blue is for shadow features that are used to help with the decision to confirm or reject.

```{r boruta plot}
lz <- lapply(1:ncol(boruta_res$ImpHistory), function(i) boruta_res$ImpHistory[is.finite(boruta_res$ImpHistory[,i]),i])

names(lz) <- colnames(boruta_res$ImpHistory)

Labels <- sort(sapply(lz,median))

plot(boruta_res, xlab = "", xaxt = "n")
axis(side = 1,las=2,labels = names(Labels), at = 1:ncol(boruta_res$ImpHistory), cex.axis = 0.7)
```

Detailed results for each variable.

```{r boruta detail}
#options(width=125)

arrange(cbind(attr=rownames(attStats(boruta_res)), attStats(boruta_res)), desc(medianImp))
```

These are the final attributes that will be used to predict house prices.

```{r}
boruta_ten <- TentativeRoughFix(boruta_res)
print(boruta_ten)

cat("\nRelevant Attributes:\n")
final_attr <- getSelectedAttributes(boruta_ten)
final_attr

final_attr <- append(final_attr, c("SalePrice.log", "Id"))

df_boruta_test <- test[final_attr]
df_boruta_train <- train[final_attr]
```

# Models
## Linear Regression (by AIC)

```{r aic model, results=FALSE}
mod1 <- lm(SalePrice.log ~ .-Id, data = df_boruta_train)

mod1_aic <- mod1 %>% step()

#variable importance sorted from highest to lowest
varImp_sort <- function(mod){
  imp <- as.data.frame(varImp(mod))
  imp <- data.frame(names   = rownames(imp),
                  overall = imp$Overall)
  imp[order(imp$overall,decreasing = T),] %>% head(20)
}

varImp_sort(mod1_aic)
```

## Linear Regression (by p-value)

```{r p model}
#####################################
# Automated model selection
# Author      : Joris Meys
# version     : 0.2
# date        : 12/01/09
#####################################
#CHANGE LOG
# 0.2   : check for empty scopevar vector
#####################################

# Function has.interaction checks whether x is part of a term in terms
# terms is a vector with names of terms from a model
has.interaction <- function(x,terms){
    out <- sapply(terms,function(i){
        sum(1-(strsplit(x,":")[[1]] %in% strsplit(i,":")[[1]]))==0
    })
    return(sum(out)>0)
}

# Function Model.select
# model is the lm object of the full model
# keep is a list of model terms to keep in the model at all times
# sig gives the significance for removal of a variable. Can be 0.1 too (see SPSS)
# verbose=T gives the F-tests, dropped var and resulting model after 
model.select <- function(model,keep,sig=0.05,verbose=F){
      counter=1
      # check input
      if(!is(model,"lm")) stop(paste(deparse(substitute(model)),"is not an lm object\n"))
      # calculate scope for drop1 function
      terms <- attr(model$terms,"term.labels")
      if(missing(keep)){ # set scopevars to all terms
          scopevars <- terms
      } else{            # select the scopevars if keep is used
          index <- match(keep,terms)
          # check if all is specified correctly
          if(sum(is.na(index))>0){
              novar <- keep[is.na(index)]
              warning(paste(
                  c(novar,"cannot be found in the model",
                  "\nThese terms are ignored in the model selection."),
                  collapse=" "))
              index <- as.vector(na.omit(index))
          }
          scopevars <- terms[-index]
      }

      # Backward model selection : 

      while(T){
          # extract the test statistics from drop.
          test <- drop1(model, scope=scopevars,test="F")

          if(verbose){
              cat("-------------STEP ",counter,"-------------\n",
              "The drop statistics : \n")
              print(test)
          }

          pval <- test[,dim(test)[2]]

          names(pval) <- rownames(test)
          pval <- sort(pval,decreasing=T)

          if(sum(is.na(pval))>0) stop(paste("Model",
              deparse(substitute(model)),"is invalid. Check if all coefficients are estimated."))

          # check if all significant
          if(pval[1]<sig) break # stops the loop if all remaining vars are sign.

          # select var to drop
          i=1
          while(T){
              dropvar <- names(pval)[i]
              check.terms <- terms[-match(dropvar,terms)]
              x <- has.interaction(dropvar,check.terms)
              if(x){i=i+1;next} else {break}              
          } # end while(T) drop var

          if(pval[i]<sig) break # stops the loop if var to remove is significant

          if(verbose){
             cat("\n--------\nTerm dropped in step",counter,":",dropvar,"\n--------\n\n")              
          }

          #update terms, scopevars and model
          scopevars <- scopevars[-match(dropvar,scopevars)]
          terms <- terms[-match(dropvar,terms)]

          formul <- as.formula(paste(".~.-",dropvar))
          model <- update(model,formul)

          if(length(scopevars)==0) {
              warning("All variables are thrown out of the model.\n",
              "No model could be specified.")
              return()
          }
          counter=counter+1
      } # end while(T) main loop
      return(model)
}

mod1_p <- model.select(mod1)
mod1_p %>% summary()
#mod1_p <- update(mod1_p,.~.-Sewer)

#view variable importance
varImp_sort(mod1_p)

compare_performance(mod1_aic,mod1_p,metrics=c("AIC","R2","RMSE"))
```

## Ensemble

All models ended up being pretty accurate with gradient boosting being the best and linear regression the worst.

```{r}
mod1 <- lm(SalePrice.log ~ .-Id-SchoolDist-Residential-Bsmt-BsmtFinType1, data=df_boruta_train)
summary(mod1)
anova(mod1, test="Chisq")

control <- trainControl(method="repeatedcv", number=10, repeats=2, savePredictions = "final", allowParallel = TRUE)
algorithm_list <- c("lm","knn","rf","gbm")

##start here. view results between using repeated cv and not

set.seed(123)
cl <- makePSOCKcluster(5)
registerDoParallel(cl)
mod <- caretList(SalePrice.log ~ .-Id-SchoolDist-Residential-Bsmt-BsmtFinType1, 
                 data=df_boruta_train, methodList=algorithm_list, trControl=control)
stopCluster(cl)

results <- resamples(mod)
summary(results)

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(results, scales=scales)
```


# Apply Model

Finally, predict results with test data frame.

```{r, warning=FALSE}
df_final <- test %>% mutate(score = predict(mod1_aic, newdata=test),
                            score = score * attr(df$SalePrice.log, 'scaled:scale') + attr(df$SalePrice.log, 'scaled:center'),
                            SalePrice = exp(score)) %>% select(Id,SalePrice)


#write_csv(df_final, "house_price_predictions.csv")
```
