---
title: "Untitled"
author: "Catherine Williams"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_fragment:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data came from [Data.world](https://data.world/inform8n/us-national-parks-visitation-1904-2016-with-boundaries). The data includes park name, location, and number of visitors.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

directory <- "09-17-19 - National Parks"

if(!getwd() == paste0("C:/Users/Cat/Google Drive/Data Analysis/Tidy Tuesday/",directory)) {
  setwd(paste0("~/Tidy Tuesday/",directory))
  }
```

# Setup

## Load Libraries

```{r library}
#devtools::install_github("hrbrmstr/albersusa")
if (!require("pacman")) install.packages("pacman")
#pacman::p_load("tidyverse","visdat","leaflet","albersusa","usmap","sf","ggthemes")
pacman::p_load("tidyverse","visdat","rgdal","maptools","ggthemes","leaflet","FRK")

theme_set(theme_classic())
```

## Import Data

```{r import}
df_parks <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv")
```

# Data Wrangling & Analysis

- There are 21,560 observations and 12 variables.

## View Data

```{r view}
glimpse(df_parks)
head(df_parks)
summary(df_parks)
sapply(df_parks, function(x) n_distinct(x)) %>% sort()

unique(df_parks$unit_type)
```

## Filter Data

I only want to analyze National Parks and do not wish to include Recreation Areas, Historical Parks, Battlefields, etc. so I will filter the data to only include "unit_type" for National Park.

This drops the number of observations down to 4,743.

```{r}
df_parks <- df_parks %>% filter(unit_type == "National Park")
```

## Missing Values

View missing values in more detail.

```{r missing}
#Visualize missing values
vis_miss(df_parks, sort_miss=TRUE)

#see count of missing values
na_values <- function(df_parks){
  na <- colSums(is.na(df_parks)) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df_parks)
```

## Data Wrangling

- Removed metadata column since it will not be used
- Filled out parkname for missing values using the unit_name column. It is just the unit_name minus the words "National Park" at the end. This will be helpful to have for visualizations.

```{r wrangle}
df_parks <- df_parks %>% select(-metadata)

df_parks$parkname <- if_else(is.na(df_parks$parkname), str_extract(df_parks$unit_name, ".+(?= National Park)"), df_parks$parkname)

# this separates Denali National Park and Denali National Preserve. Only one is needed. the numbers are identical.
df_visitors_total <- df_parks %>% filter(year == "Total") %>% mutate(parkname = fct_reorder(parkname, visitors)) %>% arrange(visitors)

df_visitors_total <- df_visitors_total[!duplicated(df_visitors_total$parkname),]

# this duplicates the vales for Denali because it is adding National Park and National Preserve together
df_visitors_annual <- df_parks %>% filter(year != "Total") %>% group_by(parkname) %>% summarize(total_visitors = sum(visitors)) %>% ungroup() %>% 
  mutate(parkname = fct_reorder(parkname, total_visitors),
         total_visitors = if_else(parkname == "Denali", total_visitors/2, total_visitors)) %>% arrange(total_visitors)

df_centroids <- read_csv('National_Park_Service__Park_Unit_Centroids.csv') %>% rename_all(tolower)
```

# Visualizations

```{r viz, out.width="100%"}

tmp_dl <- tempfile()
download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2019/STATE/tl_2019_us_state.zip", tmp_dl)
#download.file("https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_state_20m.zip", tmp_dl)

unzip(tmp_dl, exdir=tempdir())
states <- readOGR(tempdir(), "tl_2019_us_state")
#states <- readOGR(tempdir(), "cb_2018_us_state_20m")


##start here. points are not displaying right because they are in a different gps format. fix that. then see about leaflet?

remove.territories = function(.df) {
  subset(.df,
           .df$id != "MP" &
           .df$id != "GU" &
           .df$id != "PR")
}

us_aea <- spTransform(states, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))
us_aea@data$id <- rownames(us_aea@data)

alaska <- us_aea[us_aea$STUSPS=="AK",]
alaska <- elide(alaska, rotate=-30)
alaska <- elide(alaska, scale=max(apply(bbox(alaska), 1, diff)) / 2.3)
alaska <- elide(alaska, shift=c(-2100000, -2500000))
#alaska <- elide(alaska, shift=c(-119.8522102, 20.2456656))
proj4string(alaska) <- proj4string(us_aea)
alaska <- spTransform(alaska, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

hawaii <- us_aea[us_aea$STUSPS=="HI",]
hawaii <- elide(hawaii, rotate=-35)
hawaii <- elide(hawaii, shift=c(5500000, -1700000))
#hawaii <- elide(hawaii, shift=c(-47.3030715, 14.4844167))
proj4string(hawaii) <- proj4string(us_aea)
hawaii <- spTransform(hawaii, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

samoa <- us_aea[us_aea$STUSPS=="AS",]
samoa <- elide(samoa, rotate=-45)
samoa <- elide(samoa, scale=max(apply(bbox(samoa), 1, diff)) / 0.7)
samoa <- elide(samoa, shift=c(-2500000, -1500000))
#samoa <- elide(samoa, shift=c(-125.6602714, 28.0993555))
proj4string(samoa) <- proj4string(us_aea)
samoa <- spTransform(samoa, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

virgin <- us_aea[us_aea$STUSPS=="VI",]
virgin <- elide(virgin, shift=c(-1300000, 0))
#virgin <- elide(virgin, shift=c(-116.3394088, 43.8192901))
proj4string(virgin) <- proj4string(us_aea)
virgin <- spTransform(virgin, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

us_aea <- spTransform(us_aea, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
us_aea <- us_aea[!us_aea$STUSPS %in% c("AK", "HI", "AS", "VI"),]
us_aea <- rbind(us_aea, alaska, hawaii, samoa, virgin)

us50 <- fortify(us_aea, region="STUSPS")
us50 = remove.territories(us50)

df_points <- df_centroids %>% inner_join(df_visitors_annual)
df_points$id <- rownames(df_points)

### start here - trying t get the points to move the same way the boxes did
ogrDrivers()

writeOGR(df_points, "towns.kml", "towns", "KML")
points <- writeOGR(",", df_points)

points <- matrix(c(df_points$x,df_points$y), nrow=57, ncol=2)
points <- as.matrix(df_points[,1:2])

points <- Polygon(points, hole=is.logical(FALSE))

points_list <- list(points@coords)

points <- SpatialPolygons(points_list,proj4string=CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))


points <- SpatialPolygonsDataFrame(df_points,"id",c("x","y"),CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

points_ak <- points[points$state=="AK",]
points_ak <- elide(points_ak, rotate=-30)
points_ak <- elide(points_ak, scale=max(apply(bbox(points_ak), 1, diff)) / 2.3)
points_ak <- elide(points_ak, shift=c(-2100000, -2500000))
#alaska <- elide(alaska, shift=c(-119.8522102, 20.2456656))
proj4string(points_ak) <- proj4string(points)
points_ak <- spTransform(points_ak, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

#plot
ggplot() +
  geom_map(data=us50, map=us50, aes(long, lat, map_id=id, group=group), fill="white", color="dark grey", size=0.15)+
  #geom_polygon(data=us_aea2, aes(INTPTLON, INTPTLAT, group=DIVISION), fill="white", color="dark grey") +
  geom_point(data=points,aes(x,y,size = total_visitors), color="#1f4045")+
  theme_map()
```



```{r}
world <- map_data("world")

points <- df_centroids %>% inner_join(df_visitors_annual)

df_centroids %>% inner_join(df_visitors_annual) %>% 
  ggplot(aes(x,y))+
  geom_polygon(data=world, aes(long, lat, group=group), fill="white", color="#916451")+
  geom_point(aes(size = total_visitors), color="#1f4045")+
  coord_map("polyconic", xlim=c(-170,-60),ylim=c(15,75))

#################################################################
us_laea <- usa_sf("laea") %>% rename(state = iso_3166_2)

albers <- inner_join(us_laea, df_centroids) 

p <- albers %>% left_join(df_visitors_annual) %>%
  ggplot()+
  geom_sf(fill="white",color="red")+
  #geom_point(aes(x,y,size = total_visitors), color="#1f4045")+
  coord_sf(datum = NA)

p + geom_point(data=points,aes(x,y,size = total_visitors), color="#1f4045")

######################
library(sp)
us_aea2 <- spTransform(us_laea, proj4string(us))
map <- leaflet(us_aea2) 

#ggsave("national_parks.png")
```

