---
title: "Christmas Birds"
author: "Catherine Williams"
date: "June 18, 2019"
output:
  html_document:
    hightlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_depth: 2
    toc_float: yes
---
# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data came from [Bird Studies Canada](https://www.birdscanada.org/index.jsp). The data includes meteorite information such as year, bird species, and number of birds.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/../Google Drive/Data Analysis/Tidy Tuesday/06-18-19 - Christmas Bird Counts/")
```

# Setup

## Load Libraries

```{r library, message=FALSE, results=FALSE}
#function to check if packages are installed, if not then install them, and load all packages
libraries <- function(packages){
  for(package in packages){
    #checks if package is installed
    if(!require(package, character.only = TRUE)){
      #If package does not exist, then it will install
      install.packages(package, dependencies = TRUE)
      #Loads package
      library(package, character.only = TRUE)
    }
  }
}

packages <- c("data.table","tidyverse","visdat","zoo","extrafont","stringr")

libraries(packages)

theme_set(theme_classic())
```

## Import Data

```{r import, message=FALSE}
df <- fread("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-18/bird_counts.csv", stringsAsFactors = TRUE)
```
# Exploratory Data Analysis

- There are 18,706 observations and 6 variables.
- how_many_counted_by_hour is how_many_counted divided by total_hours.
- There are many NA values with total_hours which results in NAs in how_many_counted_by_hour.

## View Data

```{r view}
glimpse(df)
head(df)
summary(df)
sapply(df, function(x) n_distinct(x)) %>% sort()
```

## Missing Values

View missing values in more detail.

```{r missing}
#Visualize missing values
vis_miss(df, sort_miss=TRUE)

#see count of missing values
na_values <- function(df){
  na <- colSums(is.na(df)) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df)
```

## Visualize Trends

```{r eda, warning=FALSE, out.width="75%"}
df %>% ggplot(aes(year,(how_many_counted/total_hours)))+
  geom_bar(stat="identity")+
  labs(title="Christmas Bird Counts Per Hour Over Time", x="Year", y="Count per Hour")

df %>% ggplot(aes(total_hours, how_many_counted))+
  geom_jitter(alpha=0.3, size=1)+
  labs(title="Total Christmas Bird Counts",x="Hours", y="Count")
```

# Data Wrangling

Remove/replace missing values and drop unnecessary columns

```{r wrangle}
#remove NA hours if how_many_counted is 0
df <- df[!(df$how_many_counted==0 & is.na(df$total_hours)),]

#view missing data again
na_values(df)

# use interpolated values to replace NA values, grouped by species
df <- df %>% group_by(species) %>% mutate(total_hours = na.approx(total_hours)) %>% ungroup()

#view missing data again
na_values(df)

#drop how_many_counted_by_hour. this can be recalculated later
df <- df %>% select(1:5)
```

# Final Visualization

```{r viz, out.width="100%"}
# how many owls were seen
df_owl <- df %>% filter(str_detect(species, "Owl"))
df_owl <- df_owl %>% group_by(species) %>% summarize(counted = sum(how_many_counted))
df_owl <- df_owl %>% mutate(species = str_sub(species, end=-4))
df_owl$species <- fct_reorder(df_owl$species, df_owl$counted)

df_owl %>% ggplot(aes(species,counted))+
  geom_col(fill="steelblue")+
  geom_text(aes(label=counted), hjust=-0.1, vjust=0.5, family="Bodoni MT", fontface="bold")+
  coord_flip()+
  labs(title="Christmas Owl Spottings", subtitle="From 1921 - 2017", y="Number Spotted")+
  theme(legend.position="none",
        text=element_text(family="Bodoni MT", size=14),
        axis.title.y=element_blank(),
        plot.title=element_text(face="bold", size=24, hjust=0.175),
        plot.subtitle=element_text(hjust=0.325),
        plot.background=element_rect(fill="#fff7e6"),
        panel.background=element_rect(fill="#fff7e6"))

ggsave("christmas_owls.png", limitsize=FALSE)
```

