---
title: "Media Franchises"
author: "Catherine Williams"
date: "July 3, 2019"
output:
  html_document:
    hightlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_depth: 2
    toc_float: yes
---
# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data came from [Wikipedia](https://en.wikipedia.org/wiki/List_of_highest-grossing_media_franchises). The data includes information such as franchise name, revenue generated, year created, and owners/creators of the franchise.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
setwd("~/../Google Drive/Data Analysis/Tidy Tuesday/07-02-19 - Media Franchises/")
```

# Setup

## Load Libraries

```{r library}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse","visdat","extrafont","ggpubr","png")

theme_set(theme_classic())
```

## Import Data

```{r import}
df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-02/media_franchises.csv")
```
# Exploratory Data Analysis

- There are 321 observations and 7 variables.
- Revenue is shown in 'billions' so the numbers seem small.
- Many of the variables are characters and appear to need cleaning.
- There are no missing values.
- 1996 brought in the most revenue.
- Pokémon, Hello Kitty, Winnie the Pooh, Mickey Mouse & Friends, Star Wars, and Anpanman have earned the most revenue.
- Merchandise, Licensing & Retail is the highest earning revenue category.

## View Data

```{r view}
glimpse(df)
head(df)
summary(df)
sapply(df, function(x) n_distinct(x)) %>% sort()

df %>% count(revenue_category, sort=TRUE)
df %>% count(original_media, sort=TRUE)
```

## Missing Values

View missing values in more detail.

```{r missing}
#Visualize missing values
vis_miss(df, sort_miss=TRUE)
```

## Data Cleaning

Clean fields to make them easier to read/visualize.

```{r wrangle}
df <- df %>% mutate(original_media = case_when(original_media %in% "Book" ~ "Novel",
                                               original_media %in% "Cartoon character" ~ "Cartoon",
                                               original_media %in% "Visual novel" ~ "Comic book",
                                               original_media %in% "Animated cartoon" ~ "Animated series",
                                               TRUE ~ original_media),
                    franchise = str_replace(franchise, "A Song of Ice and Fire \\/ ", ""),
                    franchise = str_replace(franchise, "Wizarding World \\/ ", ""),
                    franchise = str_replace(franchise, "Middle-earth \\/ ", ""),
                    revenue_category = str_replace(revenue_category, ", Licensing & Retail", ""),
                    revenue_category = str_replace(revenue_category, "\\/Entertainment", ""),
                    revenue_category = str_replace(revenue_category, "\\/Games", ""),
                    revenue_category = str_replace(revenue_category, " sales", ""),
                    revenue_category = str_replace(revenue_category, "Comic or Manga", "Comic\\/Manga"))

```

## Visualizations

```{r eda}
df %>% ggplot(aes(year_created, revenue, fill=revenue_category))+
  geom_col()+
  scale_fill_viridis_d(option="E")+
  labs(title="Franchise Revenue by Year", x="Year", y="Revenue (in billions)")

df %>% filter(year_created > 1970) %>% 
  ggplot(aes(year_created, revenue, fill=revenue_category))+
  geom_col()+
  scale_fill_viridis_d(option="E")+
  labs(title="Franchise Revenue after 1970", x="Year", y="Revenue (in billions)")

df_revenue <- df %>% group_by(franchise) %>% summarize(total_revenue = sum(revenue)) %>% arrange(desc(total_revenue)) %>% ungroup()

df <- left_join(df,df_revenue)

df %>% filter(year_created == 1996) %>% mutate(franchise = fct_reorder(franchise, total_revenue)) %>%
  ggplot(aes(franchise, revenue, fill=revenue_category))+
  geom_col()+
  scale_fill_viridis_d(option="E")+
  coord_flip()+
  labs(title="Franchise Revenue in 1996", y="Revenue (in billions)")+
  theme(axis.title.y = element_blank())

df %>% filter(total_revenue > 50) %>% mutate(franchise = fct_reorder(franchise, total_revenue)) %>%
  ggplot(aes(franchise, revenue, fill=revenue_category))+
  geom_col()+
  scale_fill_viridis_d(option="E")+
  coord_flip()+
  labs(title="Top Franchises' Revenue", y="Revenue (in billions)")+
  theme(axis.title.y = element_blank())

df_revenue <- df %>% group_by(original_media) %>% summarize(media_revenue = sum(revenue)) %>% arrange(desc(media_revenue)) %>% ungroup()

df <- left_join(df,df_revenue)

df %>% mutate(original_media = fct_reorder(original_media, media_revenue)) %>%
  ggplot(aes(original_media, revenue, fill=revenue_category))+
  geom_col()+
  scale_fill_viridis_d(option="E")+
  coord_flip()+
  labs(title="Revenue by Original Media Type", x="Original Media", y="Revenue (in billions)")

df_revenue <- df %>% group_by(revenue_category) %>% summarize(category_revenue = sum(revenue)) %>% arrange(desc(category_revenue)) %>% ungroup()

df <- left_join(df,df_revenue)

df %>% mutate(revenue_category = fct_reorder(revenue_category, category_revenue)) %>%
  ggplot(aes(revenue_category, revenue, fill=original_media))+
  geom_col()+
  scale_fill_viridis_d(option="E")+
  coord_flip()+
  labs(title="Revenue by Category", x="Category", y="Revenue (in billions)")
```

# Final Visualization

Image found at http://www.greeklibrary.org/wp-content/uploads/2019/05/books.jpg

```{r viz, out.width="100%"}
img <- readPNG("books.png")

df %>% filter(original_media == "Novel") %>% mutate(franchise = fct_reorder(franchise, total_revenue)) %>%
  ggplot(aes(franchise, revenue, fill=revenue_category))+
  background_image(img)+
  geom_col(color="gray", size=0.2)+
  scale_fill_brewer(palette="Blues")+
  scale_y_continuous(expand = c(0.01,0))+
  coord_flip()+
  labs(title="Franchised Novels' Revenue by Category", y="Revenue (in billions)", fill="Category")+
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text=element_text(family="Candara", size=13),
        plot.background = element_blank(),
        panel.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_text(face="bold"),
        legend.background = element_blank(),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))

ggsave("media-franchises.png")
```

