---
title: "Meteorites"
author: "Catherine Williams"
date: "June 12, 2019"
output:
  html_document:
    hightlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_depth: 2
    toc_float: yes
---
# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data came from [NASA](https://data.nasa.gov/Space-Science/Meteorite-Landings/gh4g-9sfh/data). The data includes meteorite information such as the class, mass, and location when meteorites were found/fallen.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/../Google Drive/Data Analysis/Tidy Tuesday/06-11-19 - Meteorites")
```

# Setup

## Load Libraries

```{r library, message=FALSE, results=FALSE}
#function to check if packages are installed, if not then install them, and load all packages
libraries <- function(packages){
  for(package in packages){
    #checks if package is installed
    if(!require(package, character.only = TRUE)){
      #If package does not exist, then it will install
      install.packages(package, dependencies = TRUE)
      #Loads package
      library(package, character.only = TRUE)
    }
  }
}

packages <- c("tidyverse","readxl","visdat","modeest","maps","ggthemes","RColorBrewer","SDMTools")

libraries(packages)

theme_set(theme_classic()) #applies classic theme to all charts
```

## Import Data

```{r import, message=FALSE}
df <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")
```

# Data Wrangling

- There are 45,716 observations and 10 variables.
- There are several NA values that will be looked at in more detail later.
- name and id are both unique identifiers
- geolocation is just the summary of both lat and long
- there a bunch of coordinates entered as 0 - these are basically NA

## View Data

```{r view}
glimpse(df)
head(df)
summary(df)
sapply(df, function(x) n_distinct(x)) %>% sort()
```

## Missing Values

View missing values in more detail and remove or fix them.

```{r missing}
#Visualize missing values
vis_miss(df[,colSums(is.na(df)|df==0) >0], sort_miss=TRUE)

#see count of missing values
na_values <- function(df){
  na <- colSums(is.na(df)|df==0) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df)

#remove NA coordinates
df <- df %>% drop_na(geolocation)

#remove missing coordinates
df <- subset(df, lat!=0 & long!=0)
```

## Feature Engineering

Meteorite classification data came from [The Meteorite Market](http://www.meteoritemarket.com/type.htm). The table includes merged cells so it came over a lot cleaner after copy/pasting into Excel and then unmerging all cells.

```{r feature, message=FALSE}
#import classification table to match class and determine category
df_class <- read_xlsx("classifications.xlsx", skip=3)
df_class <- df_class %>% select(category=Category,class=`Letter Designation`,comp=`Composition Type`)

#remove rows with NA in class
df_class <- df_class %>% drop_na(class)

#set category to the correct values and clean data to make the class column optimal for matching
df_class <- df_class %>% 
  mutate(category = case_when((row_number() <= which(df_class[,2] == "R")) ~ "Chondrites",
                             (row_number() >= which(df_class[,2] == "HOW") & 
                                row_number() <= which(df_class[,2] == "WIN")) ~ "Achondrites",
                             (row_number() >= which(df_class[,2] == "H") & 
                                row_number() <= which(df_class[,2] == "D")) ~ "Irons (structural)",
                             (row_number() >= which(df_class[,2] == "IAB") & 
                                row_number() <= which(df_class[,2] == "Anom")) ~ "Irons (chemical)",
                             (row_number() >= which(df_class[,2] == "PAL")) ~ "Stony Irons",
                              TRUE ~ category),
        comp = str_replace(comp,"([:graph:]+\\s+[:graph:]+)\\s.*",""),
        class = str_split(class, ",\\s|.\\s|-")) %>% unnest(class)

#replace blank values with NAs
df_class$comp[df_class$comp == ""]<-NA

df_class <- df_class %>% mutate(comp = str_split(comp, "\\s\\(")) %>% unnest(comp)
df_class <- df_class %>% mutate(comp = str_replace(comp, "\\*|\\)", ""),
                                comp = str_extract(comp, ".*[^s$]"))

#combine the comp and class column since the data uses both variations as its classification
df_class <- df_class %>% gather(column, class, comp:class) %>% select(-column) %>% distinct() %>% drop_na()

#make sure the original df has clean class columns for matching
df <- df %>% mutate(class = str_replace_all(class, "[:punct:]+|~+", " "))

#function to find if any string value in class matches in df_class 
find_cat <- function(x, patterns, replacements=patterns, fill=NA, ...){
  stopifnot(length(patterns) == length(replacements))

  ans = rep_len(as.character(fill), length(x))    
  empty = seq_along(x)

  for(i in seq_along(patterns)) {
      greps = grepl(patterns[[i]], x[empty], ...)
      ans[empty[greps]] = replacements[[i]]  
      empty = empty[!greps]
  }
  return(ans)
}

df$category <- find_cat(df$class, df_class$class, df_class$category, NA, ignore.case = TRUE)

#view the class values that still need to be worked on
df[!df$class %in% df_class$class & df$category %in% NA,] %>% select(class,category) %>% distinct()

#classify the remaining classes
#according to https://en.wikipedia.org/wiki/Ordinary_chondrite OC is a Chondrite
df <- df %>% mutate(category = case_when(str_detect(class, "^C|LL|L|E|OC") ~ "Chondrites",
                                        TRUE ~ category),
                    category = replace_na(category, mfv(category, na.rm=T))) #the remaining values cannot be classified. Using mode to give them a value
```

## Data Types

Convert characters to factors

```{r data}
#finally change the data types back to factors and remove duplicates that were made above
df <- df %>% mutate_if(is.character,as.factor)
```

# Visualizations

## World
```{r world, out.width="100%"}
world <- map_data("world")

#practice using geom_polygon
ggplot(world, aes(x=long, y=lat, group=group))+
  geom_polygon(fill="#f2e6d9",color = "white")+
  geom_point(df, mapping=aes(group=category, color=category), alpha=0.5, size=0.5)+
  coord_map(xlim=c(-180,180))+
  theme_map()+
  scale_color_brewer(palette="Set1")+
  labs(title="Meteorite Categories Across The World", color="Category")+
  guides(color=guide_legend(override.aes=list(size=4)))+
  theme(plot.background=element_rect(fill="#ccefff"),
        legend.position=c(0, 0.09),
        legend.background=element_rect(fill="#ccefff"),
        legend.key=element_rect(fill="#ccefff"),
        plot.title=element_text(size=14, hjust=0.5),
        legend.title.align=0.5,
        legend.title=element_text(face="bold"))
```

## United States

### Prepare Data Frame
```{r usa data, message=FALSE, warning=FALSE}
state <- map_data("state")
usa <- map_data("usa")

#filter out points that do not belong in the usa
x=usa$long
y=usa$lat
long=df$long
lat=df$lat

#define the points and polygon
points <- cbind(long,lat)
polypoints <- cbind(x,y)

#plot the polygon and all points to be checked
plot(rbind(polypoints,points))
polygon(polypoints,col="red")

#check which points fall within the polygon
out <- pnt.in.poly(points,polypoints)
summary(out)

#identify points not in the polygon with an x
plot(rbind(polypoints,points))
polygon(polypoints,col='#99999990')
points(out[which(out$pip==0),1:2],pch=4,col="red")

#filter df for only the relevant points
df_usa <- bind_cols(df,out) %>% select(-long1,-lat1) %>% filter(pip==1)
```

### Visualization

```{r usa, message=FALSE, warning=FALSE, out.width="100%"}
##practice using geom_map
ggplot() +
  geom_map(data=state, aes(x=long, y=lat, group=group, map_id=region), fill="#f8f2ec",color="grey", map=state)+
  geom_map(data=usa, map=usa, aes(long, lat, map_id=region), color="black", fill=NA, size=0.5)+
  geom_point(df_usa, mapping=aes(x=long, y=lat, group=category, color=category), alpha=0.8, size=0.9)+
  coord_map("polyconic", xlim=c(-124.7, -67.1), ylim = c(25.2, 49.4)) +
  theme_map()+
  labs(title="Meteorite Categories Across The United States", color="Category")+
  scale_color_brewer(palette="Set1")+
  guides(color=guide_legend(override.aes=list(size=4)))+
  theme(plot.background=element_rect(fill="#e6f7ff"),
        legend.position=c(0.83, 0.02),
        legend.background=element_rect(fill="#e6f7ff"),
        legend.key=element_rect(fill="#e6f7ff"),
        plot.title=element_text(size=14, hjust=0.5),
        legend.title.align=0.5,
        legend.title=element_text(face="bold"))
```
