---
title: "Ramen Ratings"
author: "Catherine Williams"
date: "June 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Setup

##Load Libraries
```{r library, message=FALSE, results=FALSE}
#function to check if packages are installed, if not then install them, and load all packages
libraries <- function(packages){
  for(package in packages){
    #checks if package is installed
    if(!require(package, character.only = TRUE)){
      #If package does not exist, then it will install
      install.packages(package, dependencies = TRUE)
      #Loads package
      library(package, character.only = TRUE)
    }
  }
}

packages <- c("tidyverse","naniar","rvest","textstem","modeest","countrycode","ggridges","viridis")

libraries(packages)

theme_set(theme_classic()) #applies classic theme to all charts
```

##Import Data

```{r import, message=FALSE}
df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-04/ramen_ratings.csv")
```
#Data Wrangling
- There are 3180 observations and 6 variables.
- 1 review_number is missing while another must be duplicated, as there is 1 NA value and 3178 unique values in that column. 
- stars has 14 NA values.
- style has 2 NA values.
##View Data
```{r}
glimpse(df)
head(df)
summary(df)
sapply(df, function(x) n_distinct(x)) %>% sort()
```
##Missing Values
This visualization shows the same thing that was aleady determined before from the initial analysis. It is good to have a visual check though.

```{r missing}
#Visualize missing values
gg_miss_var(df) + labs(title="Missing Values")

#see count of NA values
df %>% is.na() %>% colSums() %>% sort(decreasing=TRUE)

#view only rows with NAs
df <- df %>% rownames_to_column()  #add row number to make it easier to locate observations
df %>% filter_all(any_vars(.=="NA"|is.na(.)))

#add in missing value for review_number
df[189:191,]
df[190,2] <- 2991

#find missing values for style and stars
df <- df %>% group_by(country, brand)%>% mutate(style = replace_na(style, mfv1(style, na.rm=T)),
                                                stars = replace_na(stars, mean(stars, na.rm=T))) %>% ungroup()

#see count of NA values again
df %>% is.na() %>% colSums() %>% sort(decreasing=TRUE)

#there are still some missing values, so we'll try again with less grouping
#find missing values for stars
df <- df %>% group_by(country)%>% mutate(stars = replace_na(stars, mean(stars, na.rm=T))) %>% ungroup()

```

##Change Data Types
```{r}
df <- df %>% mutate_if(is.character,as.factor)
```

##Data Cleaning
###Country

This uses a country list to see if country names match up with known country names. If not, they are assumed to have typos or be incorrectly labeled (such as using city names)

Country data obtained from https://data.humdata.org/dataset/countries-and-territories-beta
```{r, warning=FALSE, message=FALSE}
#import data frame with list of countries
country <- read_csv("https://docs.google.com/spreadsheets/d/1NjSI2LaS3SqbgYc0HdD8oIb7lofGtiHgoKKATCpwVdY/export?format=csv&gid=1088874596", skip=1)
head(country)

#there appear to be 2 columns with slightly different names. gathering these into one column
country <- country %>% gather(column, name, c(`#country +name +i_en +alt +v_unterm`, `#country +name +preferred`)) %>% select(name) %>% distinct()

#now sort out country values that are NOT in the country name to correct them and (hopefully) remove single, less meaningful values
country_fix <- df[!df$country %in% country$name,] %>% select(country) %>% distinct()
country_fix

#change values to appropriate names
df <- df %>% mutate(country = as.character(country),
                    country = (case_when(country %in% "USA" ~ "United States",
                                        country %in% "Dubai" ~ "United Arab Emirates",
                                        country %in% "Holland" ~ "Netherlands",
                                        country %in% "Sarawak" ~ "Malaysia",
                                        country %in% "UK" ~ "United Kingdom",
                                        country %in% "Phlippines" ~ "Philippines",
                                        TRUE ~ country)),
                    country = as.factor(country)) 
```

###Brand
```{r}
df <- df %>% mutate(brand = str_replace_all(brand, "[:punct:]+", " "),
                    brand = str_replace(brand, " s", "'s"),
                    brand = str_to_title(brand),
                    brand = str_squish(brand),
                    brand = stem_words(brand) %>% as.factor())
```
###Variety
```{r}
df <- df %>% mutate(variety = str_replace_all(variety, "[:punct:]+", " "),
                    variety = str_replace(variety, " s", "'s"),
                    variety = str_to_title(variety),
                    variety = str_squish(variety),
                    variety = stem_words(variety) %>% as.factor())
```
###Style
```{r}
df %>% count(style, sort=T)
df <- df %>% group_by(style) %>% filter(n()>4) %>% ungroup()
```

##Feature Engineering
There are a lot of countries that will be hard to visualize. A region column will be added.

```{r}
# change countries into regions
df$region <- df$country %>% countrycode(origin="country.name",destination="region") %>% as.factor()
df <- df %>% group_by(region) %>% filter(n()>10) %>% ungroup()

count(df, region, sort=TRUE)

# change countries into continents
df$continent <- df$country %>% countrycode(origin="country.name",destination="continent") %>% as.factor()

count(df, continent, sort=TRUE)
```

#Exploratory Data Analysis
##Country Ratings
```{r}
df_country <- df %>% group_by(country) %>% summarize(stars=mean(stars)) %>% ungroup()

df_country %>% mutate(country = fct_reorder(country, stars)) %>% 
  ggplot(aes(country, stars))+
  geom_bar(stat="identity")+
  coord_flip()
```

##Region Ratings
```{r clean, out.width="75%"}
df_region <- df %>% group_by(region) %>% summarize(stars=mean(stars)) %>% ungroup()

df_region %>% mutate(region = fct_reorder(region, stars)) %>%
  ggplot(aes(region, stars))+
  geom_bar(stat="identity")+
  coord_flip()
```

##Continent Ratings
```{r}
df_continent <- df %>% group_by(continent) %>% summarize(stars=mean(stars)) %>% ungroup()

df_continent %>% mutate(continent = fct_reorder(continent, stars)) %>%
  ggplot(aes(continent, stars))+
  geom_bar(stat="identity")+
  coord_flip()
```
##Style Ratings
```{r}
df_style <- df %>% group_by(style) %>% summarize(stars=mean(stars)) %>% ungroup()

df_style %>% mutate(style = fct_reorder(style, stars)) %>%
  ggplot(aes(style, stars))+
  geom_bar(stat="identity")+
  coord_flip()
```
#Final Visualization
```{r, message=FALSE}
df %>% mutate(region = fct_reorder(region, stars)) %>% 
  ggplot(aes(stars, region, fill = region))+
  geom_density_ridges(alpha=0.9) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  theme(legend.position = "none")+
  scale_x_continuous(breaks=0:5) +
  labs(title="Ramen Ratings by Region", x="Rating", y="Region")
```

