---
title: "UFO Sightings"
author: "Catherine Williams"
date: "June 26, 2019"
output:
  html_document:
    hightlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_depth: 2
    toc_float: yes
---
# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data came from [The National UFO Reporting Center](http://www.nuforc.org/). The data includes UFO sighting information such as date/time it was observed, location, UFO shape, and a description of the event.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
setwd("~/../Google Drive/Data Analysis/Tidy Tuesday/06-25-19 - UFO Sightings/")
```

# Setup

## Load Libraries

```{r library}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table","tidyverse","visdat","lubridate","rworldmap","sp","rworldxtra","maps","countrycode","ggridges")

theme_set(theme_classic())
```

## Import Data

```{r import}
df <- fread("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")
```
# Exploratory Data Analysis

- There are 80,332 observations and 11 variables.
- Most of the missing values are with state and country which could be looked up based on the coordinates.

## View Data

```{r view}
glimpse(df)
head(df)
summary(df)
sapply(df, function(x) n_distinct(x)) %>% sort()

df %>% count(ufo_shape, sort=TRUE)
```

## Missing Values

View missing values in more detail.

```{r missing}
#Visualize missing values
vis_miss(df, sort_miss=TRUE)

#see count of missing values
na_values <- function(df){
  na <- colSums(is.na(df)) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df)
```

# Data Wrangling

Remove/replace missing values and drop unnecessary columns

```{r wrangle}
df$latitude[is.na(df$latitude)] <- 33.2001

# The single argument to this function, points, is a data.frame in which:
#   - column 1 contains the longitude in degrees
#   - column 2 contains the latitude in degrees
coords2country = function(points){  
  countriesSP <- getMap(resolution='high') # uses high res map from rworldxtra

  #setting CRS directly to that from rworldmap
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  

  # use 'over' to get indices of the Polygons object containing each point 
  indices = over(pointsSP, countriesSP)

  # return the names of each country
  indices$ISO3
}

df$country <- coords2country(df[,c(11,10)])

#function to find if any string value matches, then make replacement 
find_string <- function(x, patterns, replacements=patterns, fill=NA, ...){
  stopifnot(length(patterns) == length(replacements))

  ans = rep_len(as.character(fill), length(x))    
  empty = seq_along(x)

  for(i in seq_along(patterns)) {
      greps = grepl(patterns[[i]], x[empty], ...)
      ans[empty[greps]] = replacements[[i]]  
      empty = empty[!greps]
  }
  return(ans)
}

#prepare to match on country names found in city_area column
country <- map("world")
country <- country$names %>% as_data_frame()
country$iso <- country$value %>% countrycode(origin="country.name",destination="iso3c")

country <- separate_rows(country, value, sep=":")
country <- country %>% filter(str_detect(value, "[:alpha:]+") & !is.na(iso)) %>% distinct()

df$country <- find_string(df$city_area, country$value, country$iso, df$country, ignore.case = TRUE)


#prepare to match on state abbreviations for countries in the USA
data(state.fips)
state.fips <- state.fips %>% select(abb) %>% distinct()
state.fips <- add_row(state.fips, abb = c("AK","HI"))
state.fips <- add_column(state.fips, country = "USA")

df$country <- find_string(df$state, state.fips$abb, state.fips$country, df$country, ignore.case = TRUE)

#sort out country values that are still NA
df[is.na(df$country),] %>% select(city_area) %>% distinct()

df <- df %>% 
  mutate(date_time = parse_date_time(date_time, 'mdy_hm'),
         country = as.character(country),
         country = case_when(is.na(country) & str_detect(city_area, "u\\.|uss|whitehouse") ~ "USA",
                             is.na(country) & str_detect(city_area, "viet nam") ~ "VNM", 
                             is.na(country) & str_detect(city_area, "britsh virgin islands") ~ "VGB",
                             is.na(country) & str_detect(city_area, "virgin") ~ "VIR",
                             is.na(country) & str_detect(city_area, "playa del cura|costa adeje|gibraltar") ~ "ESP",
                             is.na(country) & str_detect(city_area, "faliraki") ~ "GRC",
                             TRUE ~ country),
        ufo_shape = case_when(ufo_shape %in% "changed" ~ "changing",
                              ufo_shape %in% "round" ~ "circle",
                              ufo_shape %in% "flare" ~ "light",
                              ufo_shape %in% c("pyramid","delta") ~ "triangle",
                              is.na(ufo_shape) ~ "unknown",
                              TRUE ~ ufo_shape),
         country = replace_na(country, "Ocean"))

#view missing data again
na_values(df)


df$continent <- df$country %>% countrycode(origin="iso3c",destination="continent")
df$continent[is.na(df$continent)] <- "Other"

```

# Visualizations

```{r viz, out.width="100%"}
df_country <- df %>% group_by(country) %>% summarize(count=n()) %>% arrange(count) %>% tail(10)
df_top_country <- df %>% filter(country %in% df_country$country)


df_country %>% mutate(country = fct_reorder(country, count)) %>% 
  ggplot(aes(country, log(count), fill=count))+
  geom_col()+
  coord_flip()+
  labs(title="Top Countries for UFO Sightings")


df_top_country %>% ggplot(aes(hour(date_time),country,fill=country))+
  geom_density_ridges()+
  scale_x_continuous(limits=c(0,24),breaks=c(0,4,8,12,16,20,24),labels=c("12am","4am","8am","noon","4pm","8pm","12am"))+
  labs(title="UFO Sightings by Time of Day",x="Time",y="Country")+
  theme(legend.position="none")


df %>% ggplot(aes(as.factor(month(date_time))))+
  geom_bar(fill="olivedrab3")+
  facet_wrap(~continent, scales="free_y")+
  labs(title="UFO Sightings by Month & Continent",subtitle="From November 1906 - April 2014",x="Month",y="Number of sightings")+
  scale_x_discrete(breaks=c(1,3,5,7,9,11),labels=c("Jan","Mar","May","Jul","Sep","Nov"))+
  theme(panel.background=element_rect(fill="gray16"),
        plot.title=element_text(hjust=0.5, size=18, face="bold"),
        plot.subtitle=element_text(hjust=0.5))
              
ggsave("ufo_sightings.png")
```

