---
title: "Wildlife Impacts"
author: "Catherine Williams"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_fragment:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data came from [FAA](https://wildlife.faa.gov/). The data consists of voluntarily reported airplane wildlife strikes since 1990. The data was pre-filtered to include only the major airlines in the USA: American Airlines, Delta, Southwest, and United. The information obtained may include date, time, location, animal species, type of damage, weather conditions, airplane information, and repair costs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

directory <- "07-23-19 - Wildlife Impacts"

if(!getwd() == paste0("C:/Users/Cat/Google Drive/Data Analysis/Tidy Tuesday/",directory)) {
  setwd(paste0("~/Tidy Tuesday/",directory))
}
```

# Setup

## Load Libraries

```{r library}
if (!require("ggradar")) devtools::install_github("ricardo-bion/ggradar", dependencies = TRUE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table","tidyverse","visdat","lubridate","modeest","ggradar","scales")

theme_set(theme_classic())
```

## Import Data

```{r import}
df <- fread("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-23/wildlife_impacts.csv", stringsAsFactors=TRUE)

#remove duplicate rows
df <- df %>% distinct()
```

# Data Wrangling & Analysis

- There are 56,449 observations and 21 variables.

## View Data

```{r view}
glimpse(df)
head(df)
df %>% mutate_if(is.character, as.factor) %>% summary()
sapply(df, function(x) n_distinct(x)) %>% sort()
```

## Missing Values

View missing values in more detail.

```{r missing}
#Visualize missing values
vis_miss(df, sort_miss=TRUE, warn_large_data = FALSE)

#see count of missing values
na_values <- function(df){
  na <- colSums(is.na(df)) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df)
```

## Data Wrangling

- Dropped cost_repairs_infl_adj since the column is almost entirely missing.
- Dropped incident_month and incident_year since the information is contained in incident_date field.
- Cleaned up time by cross referencing to time_of_day and converting all times into military time.
- Merged date and time into a single column.

```{r wrangle}
#remove repair costs since there are too many missing values
df <- df %>% select(-cost_repairs_infl_adj,-incident_month,-incident_year)

#mutate date and time columns to join into a single column
df <- df %>% mutate(time = str_replace(time, "^\\d{1,2}$", paste0(time,"00")),
                    time = as.double(time),
                    time = case_when(time_of_day == "Night" & time < 1700 ~ time+1200,
                                     time_of_day == "Dusk" & time < 1200 ~ time+1200,
                                     is.na(time) ~ 0,
                                     TRUE ~ time),
                    time = str_pad(time, 4, side="left", pad="0"),
                    time = if_else(time_of_day == "Night" & time > 2400, 
                                   str_pad(time, 6, side="left", pad="0"), 
                                   str_pad(time, 6, side="right", pad="0")),
                    incident_date = str_sub(incident_date,1,10),
                    datetime = paste(incident_date,time),
                    #this will convert to date/time
                    datetime = as_datetime(datetime), 
                    #need to put it back to character in order to make proper NA replacement 
                    #it picks a random time instead of inserting 0s if this is not done
                    datetime = as.character(datetime), 
                    #replace NA values generated from illegal time formats
                    datetime = replace(datetime,is.na(datetime),incident_date[is.na(datetime)]),
                    #convert back to date/time
                    datetime = as_datetime(datetime)) %>%
  select(-time,-incident_date)

#complete time_of_day based on time of year, day, and location
df <- df %>% mutate(month = month(datetime),
                    hour = hour(datetime)) %>% 
  group_by(month,hour,state) %>% 
  mutate(time_of_day = replace_na(time_of_day, mfv1(time_of_day, na.rm=T))) %>% 
  ungroup()

na_values(df) #there are still 151 missing time_of_day values

#fill in values with educated guesses based on time of day
df <- df %>% mutate(time_of_day = as.character(time_of_day),
                    time_of_day = case_when(is.na(time_of_day) & hour > 5 & hour < 7 ~ "Dawn",
                                            is.na(time_of_day) & hour > 17 & hour < 19 ~ "Dusk",
                                            is.na(time_of_day) & hour >= 7 & hour <= 17 ~ "Day",
                                            is.na(time_of_day) & hour >=19 | hour <=5 ~ "Night",
                                            TRUE ~ time_of_day))
na_values(df)

#make airport names more readable
df <- df %>% mutate(airport = str_replace_all(airport," ARPT| INTL",""),
                    airport = str_to_title(airport) %>% as.factor())

#clean phase_of_flt
df <- df %>% mutate(phase_of_flt = str_to_lower(phase_of_flt),
                    phase_of_flt = replace_na(phase_of_flt,"unknown"),
                    phase_of_flt = str_replace(phase_of_flt, "local", "unknown"),
                    phase_of_flt = as.factor(phase_of_flt),
                    phase_of_flt = fct_relevel(phase_of_flt, c("parked","taxi","take-off run","departure","climb","en route","descent","approach","landing roll","arrival","unknown")))

summary(as.factor(df$phase_of_flt))
```

# Visualizations

The first three visualizations are misleading because there is no information on number of total flights so per capita amounts cannot be established.

The [FAA Color Scheme](https://www.color-hex.com/color-palette/10305) was used for these visualizations.

```{r viz, out.width="100%"}
#this is misrepresenting the data because there are more day and night hours in a day 
df %>% ggplot(aes(time_of_day))+
  geom_bar(fill="#0f4d92")+
  labs(title="What part of the day do most bird strikes happen?")

#this is misrepresenting the data because some times are more popular to fly than others
df %>% filter(hour != 0) %>% 
  ggplot(aes(hour, fill=time_of_day))+
  geom_bar()+
  scale_fill_manual(values=c("#eedd82","#0f4d92","#b8860b","#008000"))+
  labs(title="What time of day do most bird strikes happen?")

#this could be misrepresenting the data because some airports are busier than others
df %>% filter(airport_id != "ZZZZ") %>% count(airport) %>% filter(n > 700) %>% mutate(airport = fct_reorder(airport,n)) %>%
  ggplot(aes(airport, n))+
  geom_col(fill="#0f4d92")+
  coord_flip()+
  labs(title="Which airport has the most bird strikes?")

#final visualization
df_flt_phase <- df %>% filter(phase_of_flt != "unknown") %>% count(phase_of_flt)

df_flt_phase <- df_flt_phase %>% mutate_at(vars(n), rescale)

df_flt_phase %>% spread(phase_of_flt,n) %>% 
  ggradar(axis.label.size = 4,
          background.circle.colour = "#ffffff",
          values.radar = c("","",""),
          gridline.min.colour = "#b8860b",
          gridline.mid.colour = "#eedd82",
          gridline.max.colour = "#b8860b",
          axis.line.colour = "#008000",
          group.point.size = 4,
          group.colours = "#0f4d92",
          plot.title = "When is an airplane most likely to strike a bird?")+
  labs(caption = "Data: FAA Wildlife Strike Database | Graphic: Cat Williams @catrwilliams")+
  theme(legend.position = "none",
        plot.title = element_text(size=17, hjust=0.5, face="bold"),
        plot.caption = element_text(size=6))

ggsave("wildlife-impacts.png")
```
