---
title: "Women's World Cup"
author: "Catherine Williams"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_fragment:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

# Project Description

Tidy Tuesday has a weekly data project aimed at the R ecosystem. An emphasis is placed on understanding how to summarize and arrange data to make meaningful charts with ggplot2, tidyr, dplyr, and other tools in the tidyverse ecosystem.

# Dataset
Data comes from both [data.world](https://data.world/sportsvizsunday/womens-world-cup-data) and [Wikipedia](https://en.wikipedia.org/wiki/FIFA_Women%27s_World_Cup). The data includes information about each country such as total score, year, and whether they won, lost, or tied. There is also information about individual squads such as player name, age, position played, number of games played, and number of goals scored.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
setwd("~/../Google Drive/Data Analysis/Tidy Tuesday/07-09-19 - Womens World Cup")
```

# Setup

## Load Libraries

```{r library}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table","tidyverse","visdat")

theme_set(theme_classic())
```

## Import Data

```{r import}
df_wwc_outcomes <- fread("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/wwc_outcomes.csv")
df_squads <- fread("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/squads.csv")
df_codes <- fread("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/codes.csv")

df_wwc_outcomes <- left_join(df_wwc_outcomes, df_codes, by = "team")
```

# Exploratory Data Analysis

- There are 3 different data frames.

  - df_codes has 212 observations and 2 variables. This is just to get full country names in df_wwc_outcomes
  - df_squads has 552 observations and 9 variables.
  - df_wwc_outcomes has 568 observations and 7 variables.

## View Data

```{r view}
#df_squads
glimpse(df_squads)
head(df_squads)
summary(df_squads)

sapply(df_squads, function(x) n_distinct(x))

#df_wwc_outcomes
glimpse(df_wwc_outcomes)
head(df_wwc_outcomes)
summary(df_wwc_outcomes)

sapply(df_wwc_outcomes, function(x) n_distinct(x))
```

## Missing Values

Only df_squads has missing values- with caps and goals

```{r missing}
#Visualize missing values
vis_miss(df_squads, sort_miss=TRUE)
vis_miss(df_wwc_outcomes, sort_miss=TRUE)

#see count of missing values
na_values <- function(df){
  na <- colSums(is.na(df)) %>% sort(decreasing=TRUE)
  na[na>0]
}

na_values(df_squads)
```

## Data Wrangling

- For missing caps values, a 1 was supplied since it is assumed that player was present in the recent World Cup games.
- For missing goals values, 0 was supplied since it is assumed that if any goals were scored, they would have been recorded.

```{r wrangle}
df_squads <- df_squads %>% mutate(caps = replace_na(caps, 1),
                                  goals = replace_na(goals, 0))

df_wwc_outcomes <- df_wwc_outcomes %>% mutate(country = str_replace(country, "Ivory Coast.*", "Ivory Coast"))

# see total number of games by country
df_games <- df_wwc_outcomes %>% group_by(country) %>% summarize(games = n())

# see total number of wins and losses by country
df_wins <- df_wwc_outcomes %>% filter(win_status == "Won") %>% group_by(country) %>% summarize(wins = n())
df_loss <- df_wwc_outcomes %>% filter(win_status == "Lost") %>% group_by(country) %>% summarize(loss = n())

# get percentage of games won
df_games_won <- inner_join(df_games, df_wins) %>% mutate(percent_won = (wins/games)*100)

# get net wins/losses by country
df_games <- left_join(df_games, df_wins) %>% mutate(wins = replace_na(wins, 0))
df_games <- left_join(df_games, df_loss) %>% mutate(loss = replace_na(loss, 0),
                                                    net = wins-loss)

# capitalize country to be more consistent with FIFA France 2019 theme
df_games <- df_games %>% mutate(country = str_to_upper(country))
```

## Visualizations

```{r viz, out.width="100%"}
#Win status per country
df_wwc_outcomes %>% ggplot(aes(country))+
  geom_bar()+
  facet_wrap(~win_status)+
  coord_flip()+
  labs(title="Game Status Per Country", y="number of games")

#Win percentage per country
df_games_won %>% mutate(country = fct_reorder(country,percent_won)) %>% ggplot(aes(country, percent_won))+
  geom_col()+
  coord_flip()+
  labs(title="Percentage of Games Won Per Country", y="percent")

#ages of players
mean <- mean(df_squads$age)
median <- median(df_squads$age)

h <- hist(df_squads$age, breaks = "FD", plot = FALSE) #histogram with Freedman-Diaconis rule for binwidth

df_squads %>% ggplot(aes(age))+
  geom_histogram(aes(y = ..density..), breaks = h$breaks, alpha = 0.5, col = "white")+
  geom_vline(xintercept=mean, color="red", size=2)+
  geom_vline(xintercept=median, color="blue", size=1.5)+
  labs(title="Distribution of Players' Ages", subtitle="mean: red, median: blue")
```

# Final Visualization
Branding information was found here:
- [Colors](https://www.schemecolor.com/2019-fifa-womens-world-cup-logo-colors.php)
- [Font](http://freefootballfont.blogspot.com/2019/05/fifa-womens-world-cup-2019-france-font.html)

``` {r final}
# load custom fonts
windowsFonts(Elegance = windowsFont("Elegance"))
windowsFonts(OpenSans = windowsFont("Open Sans"))

df_games %>% mutate(country = fct_reorder(country,net)) %>% 
  ggplot(aes(country, net, fill=net < 0))+
    geom_col(width=0.95, color="white", size=0.3)+
    coord_flip()+
    scale_fill_manual(name = "net < 0", values = setNames(c("#D6000A","#F08C01"), c(T,F)))+
    scale_y_continuous(expand=c(0,0), limits=c(-37,37), breaks=c(-30,-20,-10,0,10,20,30))+
    labs(title="Women's World Cup - Net Wins Per Country", subtitle="From 1991 - 2019", y="NET WINS / LOSSES", fill="Net Game Status")+
    theme(plot.background = element_rect(fill="#23207C"),
          panel.background = element_rect(fill="#23207C"),
          legend.position="none",
          axis.line = element_line(color="white"),
          axis.text = element_text(color="white", family="OpenSans"),
          axis.text.y = element_text(size=7.5),
          axis.ticks = element_line(color="white"),
          axis.title = element_text(family="OpenSans", face="bold", color="white"),
          plot.title = element_text(hjust=0.5, family="Elegance", face="bold", size=16, color="#00B5ED"),
          plot.subtitle = element_text(hjust=0.5, family="Elegance", face="bold", size=13, color="#FDDB00"),
          axis.title.y = element_blank(),
          axis.title.x = element_text(size=8))

ggsave("womens-world-cup.png")
```
